---
title: "Executing a signed descriptor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Executing a signed descriptor}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(foceexecutorr)
```

## What this package is for

A population-PK analysis run in `sentinel-poppk` ends with a human signing a
report. At that moment the application writes a descriptor -- named after the
run (e.g. `model-run-<run-id>.json`), not literally `model.json` -- a signed,
durable description of what was decided — every candidate model as submitted,
which one was chosen and why, the fitted parameters, the accepted
data-handling, and the attestation itself.

That file is the handover. This package reads it and runs the model it names.

## Reading a descriptor

```{r}
path <- system.file("extdata", "model.json", package = "foceexecutorr")
d <- read_descriptor(path)
d
```

A descriptor is self-contained by design: an executor never calls back into the
application that produced it. That is why the file carries the *whole* submitted
tuple rather than only the winner — the fingerprint covers the tuple, and
replaying one model in isolation does not reproduce the numbers the original run
produced.

```{r}
chosen_model(d)$model_id
length(d$structural_selection$submitted_models)
```

## Running it

```{r}
res <- execute(d, backend = "inspect")
res
```

`inspect` is the only backend that ships, and it estimates nothing. That is
deliberate: the two intended backends are an approximate FOCE estimator and
NONMEM, and neither can be assumed present — NONMEM needs a licence. A stub that
returned invented parameter estimates would be indistinguishable from a real fit
to everything downstream, so the shipped default runs nothing and says so.

```{r}
res$result$executed
res$result$would_run$compartments
```

## Why `official` is not a detail

```{r}
backends()
```

Every backend declares whether its results may be treated as official, and
`execute()` reads that from the registration — never from what the backend
returns.

The reason is specific. Once two engines can both produce plausible parameter
estimates, a table of numbers no longer tells you which one you are looking at. A
fast approximate fit used to guide a decision and a qualified NONMEM run used to
support a submission are different things, and the difference is invisible at the
point where it matters most. So the executor attaches the label, from a flag the
backend had to state when it registered, and the default for anything unstated is
"not official".

## A schema v2 descriptor

sentinel-poppk has since added a second schema version: a `provenance` block
recording what built the report, and a new `declared_covariates` field on a
`structural_selection` submitted model (`covariates` there is now always
empty -- real covariate relationships only exist once the covariate_search
step has run; `declared_covariates` carries the candidates instead). Real
files are also not literally named `model.json`; sentinel-poppk names one
after its run, e.g. `model-run-<run-id>.json`.

```{r}
path2 <- system.file("extdata", "model-run-fbd67a28-06a8-40b4-9cb1-e14e6cc49ff3.json",
                      package = "foceexecutorr")
d2 <- read_descriptor(path2)
d2$schema_version

execute(d2)$result$would_run$covariates
```

`SUPPORTED_SCHEMA_VERSION` is the vector of every schema version this package
can read, so both `d` and `d2` above are valid, executable descriptors:

```{r}
SUPPORTED_SCHEMA_VERSION
```

`d2`'s `covariate_search` went on to refine its structural pick ("2cmt") into
a covariate-refined model ("cov-step3"). `chosen_model()` reports that refined
model, not the bare structural candidate:

```{r}
chosen_model(d2)$model_id
d2$structural_selection$chosen_model_id  # the bare candidate it was refined from
```

## Registering a real backend

```{r, eval = FALSE}
register_backend(
  "nonmem",
  run = function(descriptor, model, data_path, ...) {
    # compile the descriptor to a control stream, invoke NONMEM, parse results
  },
  official = TRUE,
  description = "NONMEM 7.5 via PsN"
)
```

The backend is handed the descriptor *and* the already-resolved chosen model, so
it never re-implements descriptor parsing — one reader, one place where the
schema is understood.
